---
title: "How did the Recession Hit Housing Prices in Dallas?"
output: html_document
---
```{r echo = FALSE, message = FALSE}
library(tidyverse)
library(fivethirtyeight)
library(ggthemes)
library(gridExtra)
```
```{r echo = FALSE, message = FALSE}
houses <- txhousing %>%
  filter (city == "Dallas", year > 2006, year < 2010)
beginning <- houses %>%
  filter ( year == 2007, month == 12)
end <- houses %>%
  filter (year == 2009, month == 6)
Fed_interest<- houses %>%
  filter ( year == 2007, month == 8)
Fed_further <- houses %>%
  filter ( year == 2009, month == 12)
crisis <- houses %>%
  filter ( year == 2008, month == 9)
seven <- houses %>%
  filter (year == 2007)
eight <- houses %>%
  filter (year == 2008)
nine <- houses %>%
  filter (year == 2009)
```
The price of houses and the number of houses sold in Dallas hit the highest points between May and July. Summer is usually the hunting season for houses. Therefore, the home price increases in response to high demand([Student Loan Hero](https://studentloanhero.com/featured/challenges-of-buying-home-in-summer/)). The maximum price of houses and the maximum number of houses sold each year gradually decline from the years 2007 to 2009. This illustrates that real estate industry in Dallas was influenced by the Great Recession.The data were collected from [TAMU Real Estate Center](http://recenter.tamu.edu/).
```{r echo = FALSE}
houses_long_term <- txhousing %>%
  filter (city == "Dallas", year > 2003, year < 2012)
ggplot(houses_long_term, aes(x = date, y = sales)) +
  geom_line() +
  annotate("rect", fill = "#fbb4ae", alpha = 0.6, 
        xmin = 2007, xmax = 2009,
        ymin = -Inf, ymax = Inf) +
  labs(x = "Year", y = "Number of Sales", title = "FIGURE 1 : Number of Houses Sold in Dallas from 2004 to 2012", caption = "Source : TAMU Real Estate Center") +
  scale_x_continuous(breaks = c(2004.000,2005.000, 2006.000, 2007.000,2008.000, 2009.000, 2010.000, 2011.000, 2012.000),
                         labels = c(2004, 2005,2006,2007,2008, 2009, 2010, 2011, 2012))+
  theme_economist()+
  annotate("text", x = 2008, y = 6500, label = "Great Recession", size = 5, color = "darkred")
```
Figure shows trendline of the number of houses sold between 2004 and 2012. We highlighted the timeline of the Great Recession red allow the audience to focus on the sharp decline in housing sales in Dallas during that time. 
```{r echo = FALSE, message = FALSE}
houses = houses %>%
  group_by(year) %>%
  mutate(highlight = coef(lm(volume/(sales*10000) ~ factor(month)))[2],
  highlight = ifelse(highlight > 0,
                     ifelse(volume/(sales*10000) == max(volume/(sales*10000)),"Max of the Year", "Others")))
ggplot(houses, aes(x = factor(month), y = volume/(sales*10000), color = highlight)) +
  geom_vline(data= beginning, aes(xintercept = month),col= "#fbb4ae" , size = 3, alpha = 0.6)+
     geom_vline(data= end, aes(xintercept= month), col= "#fbb4ae" , size = 3, alpha = 0.6)+
     geom_vline(data= Fed_interest, aes(xintercept = month), col= "#5aae61" , size = 3, alpha = 0.6)+
     geom_vline(data= Fed_further, aes(xintercept = month), col= "#5aae61" , size = 3, alpha = 0.6)+
 geom_vline(data= crisis, aes(xintercept = month), col= "#fee08b" , size = 3, alpha = 0.6)+
 geom_line(aes(x = month, y = volume/(sales*10000)),color = "#762a83") +
 geom_point(size = 2) +
 facet_wrap (~ year) +
 labs(x = "Months", y = "Price of the houses (in 10,000 $)", title = "FIGURE 2 : Prices of House in Dallas between 2007-2009", caption = "Source : TAMU Real Estate Center", subtitle = "Green Line = FED lowered interest rate / Orange = The beginning and end of Great Recession
      / Yellow = Global Financial Crisis") +
  scale_x_discrete(label = month.abb) +
 theme_economist() +
 theme(axis.text.x=element_text(angle=90 ,hjust=1,vjust=0.5)) 
```
```{r, echo =FALSE}
houses = houses %>%
  group_by(year) %>%
  mutate(highlight = coef(lm(sales ~ factor(month)))[2],
  highlight = ifelse(highlight > 0,
                     ifelse(sales == max(sales),"Max of the Year","Others")))
ggplot(houses, aes(x = factor(month), y = sales/1000, color = highlight)) +
  geom_vline(data= beginning, aes(xintercept = month),col= "#fbb4ae" , size = 3, alpha = 0.6)+
     geom_vline(data= end, aes(xintercept= month), col= "#fbb4ae" , size = 3, alpha = 0.6)+
     geom_vline(data= Fed_interest, aes(xintercept = month), col= "#5aae61" , size = 3, alpha = 0.6)+
     geom_vline(data= Fed_further, aes(xintercept = month), col= "#5aae61" , size = 3, alpha = 0.6)+
 geom_vline(data= crisis, aes(xintercept = month), col= "#fee08b" , size = 3, alpha = 0.6)+
 geom_line(aes(x = month, y = sales/1000),color = "#1b7837") +
 geom_point(size = 2) +
 facet_wrap(~ year) +
 labs(x = "Months", y = "Number of Houses Sold (in 1000)", title = "FIGURE 3 : Numbers of Houses Sold in Dallas between 2007-2009", caption = "Source : TAMU Real Estate Center", subtitle = "Green Line = FED lowered interest rate / Orange = The beginning and end of Great Recession/
              Yellow = Global Financial Crisis") +
  scale_x_discrete ( label = month.abb) +
 theme_economist()  +
 theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
``` 
<br />We chose to focus on the data between 2007 and 2009 since the period includes the recession as well as the beginning of the recovery period. Figure 2 focuses on the price of houses while Figure 3 focuses on the number of houses sold. We used line graphs ( using lines and points) to show the trend over these three years. We have changed the point color to emphasize the maximum price and the number of houses sold for each year to draw the user's attention. The green vertical lines on the first two plots indicate the time when Federal Reserve reduces the federal fund rate. The orange vertical lines indicate the beginning and the end of the Great Recession defined by the National Bureau of Economic Research. The yellow vertical line indicates the year when Lehman Brothers, a global financial service firm, bankrupt, which is said to be the largest bankruptcy in the U.S. history. The collapse of the top bank in the world triggered Global Panic.  